The Spatial and Temporal Domains of Modern Ecology

### 2018-02-01 paper/main.tex
First cut edits to shape paper to NEE formatting for publications (methods last, but reviewers' final suggestions).

@@

\baselineskip24pt

@@

\section*{Abstract}
\textbf{To properly understand ecological phenomena, it is necessary to observe them across a range of spatial and temporal scales. Ecologists first raised this point in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly. To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 observational studies published between 2004-2014. We found that the scale domains of observations are fairly narrow, and are collected primarily with conventional field techniques. In the spatial domain, most observations have resolutions of $\leq$1 m2 and extents of $\leq$10,000 ha. In the temporal domain, most observations were either unreplicated or of low frequency ($\geq$1 month interval), and were made over relatively short durations ($\leq$1 year). Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger. Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing portions of space and time not truly captured by replicates and raising concerns about observational comprehensiveness. Adding to these concerns, scales were not clearly reported in most studies, suggesting that it is a minor consideration. Journals can help mitigate this problem by implementing scale reporting standards, which can spur ecologists to more rapidly adopt new observational technologies, and thereby close key gaps in current observational domains.}

\linenumbers
\vspace{10pt}
The scales at which ecosystems are observed play a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}. This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.

@@

Our analysis focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1, and see SI for full definitions). Resolution is the area of an individual spatial replicate within which a complete measurement (as opposed to a sub-sample) of the feature of interest was made. Extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g. forest patches in grassland habitats), the summed area of sampled patches. Interval refers to the average time elapsed between individual temporal replicates. Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation. We also assessed observational scales within two additional dimensions, \emph{actual} extent (the summed area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates). We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e. how much space and time is covered by the measurement) differ from the scales that the observations are explicitly or implicitly intended to represent. This difference may contain important information about how effectively ecological observations characterize ecological phenomena. First, an increasing gap between actual and intended observational scale inherently implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data. Second (and relatedly), since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap may increase the likelihood of encountering unanticipated data challenges such as censored data (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between observations.

@@

%We extracted these scale dimensions from 379 observations that were reported in 134 papers that were published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. 
%

Our analysis was based on a review of 348 papers randomly selected from 42,918 published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. We extracted scale data from 378 observations of ``natural'' (i.e. non-experimentally manipulated) ecological features that were reported within 133 of the reviewed papers (plus an additional 62 that these cited as the source of observations). We excluded experiments because they tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}, and would therefore likely bias our findings towards finer scales, while minimizing the impact that new observing methods (e.g. satellite imaging, wireless sensing) may have had in expanding the scales of ecological investigation \cite{turner_remote_2003,pettorelli_satellite_2014,porter_wireless_2005}.

To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI). We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig. 1). We constructed kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four primary (resolution, extent, interval, duration) space-time dimensions (Fig. 2).

\noindent \textbf{Results}\\
\noindent \textbf{\emph{Observational methods}}\\
To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data (hereafter remote observations), and paleo-reconstruction approaches. Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.9\%, and paleo-reconstruction for less than 0.8\%.

\noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $<$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $\geq$1 ha (Fig. 1A). These distributions primarily reflect those of field observations, the dominant observational methodology. Examining the distributions for each observational method (Fig. S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).

@@

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals. With respect to temporally replicated observations (Fig. S1 in SI), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).

@@

\noindent \textbf{\emph{Spatial and temporal domains}}\\
Juxtaposing these observational dimensions provides further insight into the spatial and temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident. The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Fig. S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1000 m$^2$) but finer median interval ($\sim$1 day). Paleo-reconstructions and automated sensing techniques were both finely resolved (median between10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.

@@

Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging from $\sim$1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $<$10 ha extent. These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig. S2 in SI) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.

@@

\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover. To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation. We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig. 3). Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8.3 orders of magnitude between 100 million and 1 billion ha of extent (Fig. 3A). The difference fell to 3 orders of magnitude between 1-10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing. Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Fig. S3 in SI for method-specific plots).

The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig. S3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records. The mean difference magnitudes for the other three observing methods ranged from just over 3 (field observations and automated sensing) to nearly 6 (remote observations).

\noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains. The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63\%, 60\%, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively). The inevitable estimation errors may have biased our overall findings. However, we attempted to quantify and account for this error by assessing inter-rater disagreement and incorporating this uncertainty into our resampling methodology. The resulting confidence intervals (Fig. 1) suggest that it was unlikely that estimation errors unduly influenced our findings.

Another potential source of bias lies within our scale-estimation protocols, chiefly with respect to our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement). We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken. For these, our estimates are finer than what the studies' authors apparently considered to be the resolution of spatial replicates. Our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations. For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}. Additionally, the token 1 second (SI) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) mean that we underestimated actual duration-duration differences for this observational method (Fig. S3). However, the relatively low proportion of remote observations suggests that the impact on our overall findings was negligible (Fig. 3).

The relatively small sample size of our study may also bias these findings. Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period. Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method. The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g. Fig. 2B, C; however, the interval and duration values are probably more representative).

@@

\noindent \textbf{Discussion}\\
\noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
Our results suggest that the scale domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques. Spatially, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B). In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig. 1C) and relatively short duration ($\leq$1 year; Fig. 1D).

Contrasting observational dimensions reveals that larger extents are associated with larger spatial replicates (Fig. 2C), while longer durations tend to be associated with longer intervals (Fig. 2B). The latter association primarily reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation, the dominant observational mode, but is also reflected by the relative domain locations of the four main observational methods, suggesting that high observational frequency poses a durational cost across observational methods. A similar cost tradeoff is also evident in the inverse relationship between resolution and interval that dominates that domain space (Fig. 2A), which again primarily relates to characteristics of field-based observations, in which larger spatial replicates demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}. Less obvious is the opposite tradeoff that occurs with remote observation (but shown in Fig. S2 in SI), where finer resolution is desired to increase detail, but typically comes at the cost of longer intervals \cite{estes_platform_2016}.

@@

Have these domains changed since the seminal papers on scale first began to appear in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}? A comprehensive answer to this question would require a similar study focused on earlier literature, but an analysis of results presented in three prior studies provides partial insight. The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}. The average duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, compared to 3.3 years for observations in our sample (or 5.1 if temporally un-replicated observations are excluded). The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986. The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,496,070 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments. The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}). The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs. Extent in our sample was substantially larger according to multiple summary statistics, including the mean (114,965,072 ha in our study versus 368,403 ha), median (5,051 ha versus 9 ha), and 95th percentile (46,424,808 ha versus 136,000 ha).

@@

However, even though observational extent appears to be increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates purportedly represent (Fig. 3A). A similarly large discrepancy is also evident between the time spans covered by repeated observations and the time that is spent actually observing a phenomenon (Fig. 3B). These differences between the actual and \emph{representative} scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}. Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (as an example with respect to temporal replication, annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}. To capture such behavior, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.

@@

\noindent \textbf{\emph{Looking forward}}\\
Our study suggests that the concept of scale has yet to fully permeate the discipline of ecology. Evidence for this assertion lies in the continued narrowness of ecology's observational scale domains and the poor documentation of scale dimensions in the literature. However, the increasing extent of ecological observations, enabled by remote sensing and presumably motivated by many ecologists' appreciation of scale-related issues, suggests that ecology's scale domains are gradually changing. In the coming years, the accelerating gains in technology and analytical methods will allow researchers new and unprecedented capabilities to peer into, and thus close, the prominent holes in observational scale domains. A renewed, discipline-wide focus on scale's importance, including the adoption of stricter scale-reporting standards by journals, will help spur ecologists to address these gaps, while fostering the improved transferability of knowledge within the discipline.

>>>>

\usepackage[
  breaklinks=true,
  colorlinks=true,
  linkcolor=blue,anchorcolor=blue,
  citecolor=blue,filecolor=blue,
  menucolor=blue,pagecolor=blue,
  urlcolor=blue]{hyperref}

@@

\baselineskip12pt

@@

\section*{Abstract}
\textbf{To understand ecological phenomena, it is necessary to observe their behavior across multiple spatial and temporal scales. This need was first highlighted in the 1980s, and since then technology has opened previously inaccessible scales to observation. Have there been corresponding changes in the scales observed by modern ecologists? To help answer this question, we analyzed the resolution, extent, interval, and duration of observations (excluding experiments) in 348 studies published between 2004-2014. We found that observational scales were generally narrow because ecologists still primarily use conventional field techniques. In the spatial domain, most observations had resolutions $\leq$1 m$^2$ and extents $\leq$10,000 ha. In the temporal domain, most observations were either unreplicated or infrequently repeated ($\geq$1 month interval), and $\leq$1 year in duration. Compared to studies conducted prior to 2004, observational durations and resolutions appear largely unchanged, but intervals have become finer and extents larger. We also found a large gulf between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising concerns about observational comprehensiveness. Furthermore, most studies did not clearly report scale, suggesting that it remains a minor concern. We suggest that ecologists can obtain better understanding of the scales represented by observations by incorporating autocorrelation measures, while journals can promote attentiveness to scale by implementing scale reporting standards.}

%To fully understand ecological phenomena, we must observe them across multiple spatial and temporal scales. Beginning in the 1980s, several prominent ecological studies to this need beginning in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly. To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 observational studies published between 2004-2014. We found that most observations are collected with conventional field techniques, resulting in generally narrow observational scale domains. In the spatial domain, most observations have resolutions of $\leq$1 m$^2$ and extents of $\leq$10,000 ha. In the temporal domain, most observations were either unreplicated or of low frequency ($\geq$1 month interval), and had relatively short durations ($\leq$1 year). Compared to studies conducted prior to the early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger. Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing portions of space and time not truly captured by replicates, which raises concerns about observational comprehensiveness. Adding to these concerns, scales were not clearly reported in most studies, suggesting that scale is a minor consideration. Journals can help mitigate this problem by implementing scale reporting standards, which can spur ecologists to more rapidly adopt new observational technologies, and thereby close key gaps in current observational domains.}

%\linenumbers
\vspace{10pt}
\noindent \textbf{Introduction}
\vspace{5pt}
\\
The scales at which ecosystems are observed play a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.  Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}. This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.

@@

Our study focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1, and see Methods for full definitions). We also assessed observational scales within two additional dimensions, \emph{actual} extent (the summed area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates). We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e. how much space and time is covered by the measurement) differ from the scales that the observations ostensibly represent. This difference may contain important information about how effectively ecological observations characterize ecological phenomena. First, an increasing gap between actual and ostensible observational scale implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data. Second (and relatedly), since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap may increase the likelihood of encountering unanticipated data challenges such as censoring (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between observations.

@@

We undertook three analyses to characterize the scale domains of modern ecological observations. We first evaluated the distribution of observations (domain) within each of the four primary dimensions. Next, we examined observational domains within four different two-dimensional juxtapostions (interval versus resolution; interval versus duration; extent versus resolution; extent versus duration). Lastly, we quantified the differences between actual and ostensible scales of observation (i.e. between actual extent and extent, and actual duration and duration), and examined how these differences varied within the domain of the corresponding ostensible scale.

\vspace{10pt}
\noindent \textbf{Results}
\vspace{5pt}
\\
We reviewed 348 papers randomly selected from 42,918 titles published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. We extracted and analyzed scale data from 378 observations of ``natural'' (i.e. non-experimentally manipulated) ecological features that were reported within 133 of the reviewed papers (plus an additional 62 that these cited as the source of observations). The majority of sampled observations were collected using manual field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).

\vspace{5pt}
\noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $<$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $\geq$1 ha (Fig. 1A). These distributions primarily reflect those of field observations, the dominant observational methodology. Examining the distributions for each observational method (Supplementary Figure 1) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).

@@

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals. With respect to temporally replicated observations (Supplementary Figure 1), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).

@@

\vspace{5pt}
\noindent \textbf{\emph{Spatial and temporal domains}}\\
Juxtaposing these observational dimensions provides further insight into the spatial and temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident. The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Supplementary Figure 2): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day). Paleo-reconstructions and automated sensing techniques were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.

@@

Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging from $\sim$1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $<$10 ha extent. These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.

@@

\vspace{5pt}
\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover. To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation. We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig. 3). Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8.3 orders of magnitude between 100 million and 1 billion ha of extent (Fig. 3A). The difference fell to 3 orders of magnitude between 1-10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing. Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Supplementary Figure 3 for method-specific plots).

The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig. 3B). As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Supplementary Figure 3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records. The mean difference magnitudes for the other three observing methods ranged from just over 3 (field observations and automated sensing) to nearly 6 (remote observations).

\vspace{5pt}
\noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains. The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63\%, 60\%, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively). The inevitable estimation errors may have biased our findings. However, we attempted to quantify and account for this error by assessing between-observer variability and incorporating this uncertainty into our resampling methodology (Supplementary Results). The resulting confidence intervals (Fig. 1) suggest that it was unlikely that estimation errors unduly influenced our findings.

Another potential source of bias lies within our scale-estimation protocols, chiefly with respect to our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement). We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken. For these, our estimates are finer than what the studies' authors apparently considered to be the resolution of spatial replicates. Our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations. For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}. Additionally, the token 1 second (Supplemental Results) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) mean that we underestimated actual duration-duration differences for this observational method (Supplementary Figure 3). However, the relatively low proportion of remote observations suggests that the impact on our overall findings was negligible (Fig. 3).

It is also possible that our findings misrepresent the characteristic scales of ecological observations, due to sampling error. Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period. Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method. The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g. Fig. 2B, C; however, the interval and duration values are probably more representative).

@@

\vspace{10pt}
\noindent \textbf{Discussion}
\vspace{5pt}
\\
\noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
Our results suggest that the scale domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques. Spatially, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B). In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig. 1C) and relatively short duration ($\leq$1 year; Fig. 1D).

Contrasting observational dimensions reveals that larger extents are associated with larger spatial replicates (Fig. 2C), while longer durations tend to be associated with longer intervals (Fig. 2B). The latter association primarily reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation, the dominant observational mode, but is also reflected by the relative domain locations of the four main observational methods, suggesting that high observational frequency poses a durational cost across observational methods. A similar cost tradeoff is also evident in the inverse relationship between resolution and interval that dominates that domain space (Fig. 2A), which again primarily relates to characteristics of field-based observations, in which larger spatial replicates demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}. Less obvious is the opposite tradeoff that occurs with remote observation (Supplementary Figure 2), where finer resolution is desired to increase detail, but typically comes at the cost of longer intervals \cite{estes_platform_2016}.

@@

Have these domains changed since the seminal papers on scale first began to appear in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}? A comprehensive answer to this question would require a similar study focused on earlier literature, but an analysis of results presented in three prior studies provides partial insight. The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}. The mean duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, versus 3.3 years in our sample (or 5.1 if temporally un-replicated observations are excluded). The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who extracted resolutions from 97 community ecology experiments published in \emph{Ecology} between 1980-1986. The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments. The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}). The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs. Extent in our sample was substantially larger according to multiple summary statistics, including the mean (368,403 ha versus 114,965,072 ha in our study), median (9 ha versus 5,051 ha), and 95th percentile (136,000 ha versus 46,424,808 ha; this value is smaller than the mean, which is skewed by a small number of very large extent observations).

@@

However, even though observational extent appears to be increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates ostensibly represent (Fig. 3A). A similarly large discrepancy is also evident between the time spans covered by repeated observations and the time that is spent actually observing a phenomenon (Fig. 3B). These differences between the actual and ostensible scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}. Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (as an example with respect to temporal replication, annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena. For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}. To capture such behavior, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.

@@

\vspace{5pt}
\noindent \textbf{\emph{Looking forward}}\\
Our study suggests that the concept of scale has yet to fully permeate the discipline of ecology. Evidence for this assertion lies in the continued narrowness of ecology's observational scale domains and the poor documentation of scale dimensions in the literature. However, the increasing extent of ecological observations, enabled by remote sensing and presumably motivated by many ecologists' appreciation of scale-related issues, suggests that ecology's scale domains are gradually changing. In the coming years, the accelerating gains in technology and analytical methods will allow researchers new and unprecedented capabilities to peer into, and thus close, the prominent holes in observational scale domains. A renewed, discipline-wide focus on scale's importance, including the adoption of stricter scale-reporting standards by journals, will help spur ecologists to address these gaps, while fostering the improved transferability of knowledge within the discipline.
